---
title: "RWorksheet_Perez#4a"
output: pdf_document
date: "2024-10-14"
---

1. The table below shows the data about shoe size and height. Create a data frame.

```{r}
shoe_data <- read.csv("/cloud/project/ShoeData.csv")
shoe_data
```
a. Describe the data.

The data consists of the respondents' shoe sizes, as well as their height and gender. 

b. Create a subset by males and females with their corresponding shoe size and height.
What its result? Show the R scripts.

```{r}
males <- subset(shoe_data, Gender == "M", select = c(Shoe.size, Height))
males 
```
```{r}
females <- subset(shoe_data, Gender == "F", select = c(Shoe.size, Height))
females
```

c. Find the mean of shoe size and height of the respondents. Write the R scripts and its result.

```{r}
mean(shoe_data$Shoe.size)
```
```{r}
mean(shoe_data$Height)
```

d. Is there a relationship between shoe size and height? Why?

```{r}
relationship <- cor(shoe_data$Height, shoe_data$Shoe.size)
relationship
```
The results show that there is some kind of relationship between shoe size and height, but it's not the strongest. Some shoe sizes go up with height, but a few doesn't seem to be the case.

2. Construct character vector months to a factor with factor() and assign the result to factor_months_vector. Print out factor_months_vector and assert that R prints out the factor levels below the actual values.

Consider data consisting of the names of months:
"March","April","January","November","January",
"September","October","September","November","August",
"January","November","November","February","May","August", "July","December","August","August","September","November","February, April"
 
```{r}
months_vector <- c("March","April","January","November","January",
"September","October","September","November","August",
"January","November","November","February","May","August", "July","December","August","August","September","November","February, April")
factor_months_vector <- factor(months_vector)
factor_months_vector
```

3. Then check the summary() of the months_vector and factor_months_vector. | Inter-
pret the results of both vectors. Are they both equally useful in this case?
```{r}
summary(months_vector)
```

```{r}
summary(factor_months_vector)
```

4. Create a vector and factor for the table below.

Direction Frequency
East 1
West 4
North 3

```{r}
direction <- c("East", "West", "North")
frequency <- c(1, 4, 3)
factor_direction <- factor(direction)
factor_frequency <- factor(frequency)
```
```{r}
factor_direction
```
```{r}
factor_frequency
```

Note: Apply the factor function with required order of the level.
new_order_data <- factor(factor_data,levels = c("East","West","North"))
print(new_order_data)

```{r}
new_order_direction <- factor(factor_direction, levels = c("East", "West", "North"))
print(new_order_direction)
```
```{r}
new_order_frequency <- factor(factor_frequency, levels = c(1, 4, 3))
print(new_order_frequency)
```

5. Enter the data below in Excel with file name = import_march.csv
```{r}
read.csv("/cloud/project/import_march.csv")
```

a. Import the excel file into the Environment Pane using read.table() function.
Write the code.

```{r}
march <- read.table("/cloud/project/import_march.csv", header = TRUE, sep = ",")
```

b. View the dataset. Write the R scripts and its result.

```{r}
march
```