---
title: "RWorksheet_Perez#4b"
output:
  pdf_document: default
date: "2024-11-04"
---

---
title: "RWorksheet_Perez#4b"
output: pdf_document
date: "2024-10-28"
---

Using Loop Function

for() loop

1. Using the for loop, create an R script that will display a 5x5 matrix as shown in
Figure 1. It must contain vectorA = [1,2,3,4,5] and a 5 x 5 zero matrix.
Hint: Use abs() function to get the absolute value

```{r}
vectorA <- 1:5
matrixA <- matrix(0, nrow=5, ncol=5)
for (i in 1:5) {
    matrixA[i, ] <- abs(vectorA - i)
}
print(matrixA)
```
2. Print the string "*" using for() function. The output should be the same as shown in Figure. 

```{r}
for (i in 1:5) {
    cat(rep("*", i), "\n")
}
```

3. Get an input from the user to print the Fibonacci sequence starting from the 1st input up to 500. Use repeat and break statements. Write the R Scripts and its output.


```{r}
first <- as.integer(readline(prompt="Enter the first number of the Fibonacci sequence: "))
second <- as.integer(readline(prompt="Enter the second number of the Fibonacci sequence: "))

#for knitting purposes
if (is.null(first) || is.na(first)) first <- 1
if (is.null(second)|| is.na(second)) second <- 1

fibonacci <- c(first, second)
prev <- first
current <- second
repeat {
  next_val <- prev + current
  if (next_val > 500) break
      fibonacci <- c(fibonacci, next_val)
      prev <- current
      current <- next_val
}
print(fibonacci)
```
Using Basic Graphics (plot(),barplot(),pie(),hist())
4. Import the dataset as shown in Figure 1 you have created previously.
```{r}
shoe_data <- read.csv("/cloud/project/ShoeData.csv")
shoe_data
```

a. What is the R script for importing an excel or a csv file? Display the first 6 rows of the dataset? Show your codes and its result. 
```{r}
shoe_data <- read.csv("/cloud/project/ShoeData.csv")
head(shoe_data)
```

b. Create a subset for gender(female and male). How many observations are there in
Male? How about in Female? Write the R scripts and its output.

```{r}
female <- subset(shoe_data, Gender == "F")
male <- subset(shoe_data, Gender == "M")
cat("Female count:", nrow(female),"\n")
cat("Male count:", nrow(male),"\n")
```
c. Create a graph for the number of males and females for Household Data. Use plot(), chart type = barplot. Make sure to place title, legends, and colors. Write the R scripts and its result.
```{r}
gender_count <- table(shoe_data$Gender)
barplot(gender_count, main= "Gender Distribution", col = c("pink", "green"),
xlab="Gender", ylab="Count", legend=TRUE)
```

5. The monthly income of Dela Cruz family was spent on the following:

Food Electricity Savings Miscellaneous

60 10 5 25

a. Create a piechart that will include labels in percentage.Add some colors and title of the chart. Write the R scripts and show its output.

```{r}
install.packages("ggplot2")
library(ggplot2)

categories <- c("Food", "Electricity", "Savings", "Miscellaneous")
spent <- c(60, 10, 5, 25)
amount_spent <- data.frame(categories, spent)
amount_spent$percentage <- amount_spent$spent / sum(amount_spent$spent) * 100
ggplot(amount_spent, aes(x = "", y = percentage, fill = categories)) + geom_bar(stat = "identity", width = 1) + coord_polar("y") +
geom_text(aes(label = paste0(round(percentage), "%")), position = position_stack(vjust = 0.5)) + labs(title = "Monthly Income Distribution of Dela Cruz Family") + theme_void() + scale_fill_brewer(palette = "Set2")
```

6. Use the iris dataset.

a. Check for the structure of the dataset using the str() function. Describe what you have seen in the output.

```{r}
data(iris)
str(iris)
```
b. Create an R object that will contain the mean of the sepal.length,
sepal.width,petal.length,and petal.width. What is the R script and its result?

```{r}
mean <- colMeans(iris[, 1:4])
mean
```
c. Create a pie chart for the Species distribution. Add title, legends, and colors. Write the R script and its result.
```{r}
specs <- table(iris$Species)
pie(specs,
main = "Species Distribution in Iris Dataset",
col = c("lightblue", "violet", "lightpink"),
labels = paste(names(specs), specs))
legend("topright", legend = names(specs), fill = c("lightblue", "violet", "lightpink"))

```
d. Subset the species into setosa, versicolor, and virginica. Write the R scripts
and show the last six (6) rows of each species.

```{r}
setosa <- iris[iris$Species == "setosa", ]
versicolor <- iris[iris$Species == "versicolor", ]
virginica <- iris[iris$Species == "virginica", ]

tail(setosa)
tail(versicolor)
tail(virginica)
```
e. Create a scatterplot of the sepal.length and sepal.width using the different
species(setosa,versicolor,virginica). Add a title = “Iris Dataset”, subtitle = “Sepal width and length, labels for the x and y axis, the pch symbol and colors should be based on the species.

Hint: Need to convert to factors the species to store categorical variables.

```{r}
iris$Species <- as.factor(iris$Species)

plot(iris$Sepal.Length, iris$Sepal.Width,
col = iris$Species,
pch = 19,
main = "Iris Dataset",
sub = "Sepal Width and Length",
xlab = "Sepal Length",
ylab = "Sepal Width")
legend("topright", legend = levels(iris$Species), col = 1:3, pch = 19)
```
f. Interpret the result.

The three species show a clear distinction based on Sepal Length and Width, with minimal overlap. Setosa is easily identifiable due to its shorter Sepal Length, while Versicolor and Virginica have some overlap but can still be differentiated, as Virginica typically has longer Sepal Lengths.

7. Import the alexa-file.xlsx. Check on the variations. Notice that there are ex-
tra whitespaces among black variants (Black Dot, Black Plus, Black Show, Black
Spot). Also on the white variants (White Dot, White Plus, White Show, White Spot).
```{r}
install.packages("readxl")
library(readxl)
alexa <- read_xlsx("/cloud/project/alexa_file.xlsx")
```

a. Rename the white and black variants by using gsub() function.

```{r}
alexa$variation <- gsub("Black\\s+", "Black", alexa$variation)
alexa$variation <- gsub("White\\s+", "White", alexa$variation)
```

b. Get the total number of each variations and save it into another object. Save the object as variations.RData. Write the R scripts. What is its result?
Hint: Use the dplyr package. Make sure to install it before loading the package.

```{r}
install.packages("dplyr")
library(dplyr)
variationCount <- alexa %>% count(variation)
save(variationCount, file = "variations.RData")
```

c. From the variations.RData, create a barplot(). Complete the details of the chart
which include the title, color, labels of each bar.

```{r}
load("variations.RData")
barplot(variationCount$n, names.arg = variationCount$variation, col = "skyblue", main = "Alexa Variant Distribution", xlab = "Variant", ylab = "Count")

```

d. Create a barplot() for the black and white variations. Plot it in 1 frame, side by side. Complete the details of the chart.

```{r}
variantCount <- alexa %>%
  group_by(variation) %>%
  summarize(count = n())
```

```{r}
black_variants <- variantCount %>% 
  filter(grepl("Black", variation))
white_variants <- variantCount %>% 
  filter(grepl("White", variation))

print(black_variants)
print(white_variants)
```

```{r}
black_plot <- ggplot(black_variants, aes(x = variation, y = count, fill = variation)) +
  geom_bar(stat = "identity") +
  labs(title = "Black Variants", x = "Values", y = "Total Numbers") +
  theme_minimal() +
  theme(legend.position = "none") +
  scale_fill_manual(values = c("black", "red", "green", "blue", "cyan"))
```

```{r}
white_plot <- ggplot(white_variants, aes(x = variation, y = count, fill = variation)) +
  geom_bar(stat = "identity") +
  labs(title = "White Variants", x = "Values", y = "Total Numbers") +
  theme_minimal() +
  theme(legend.position = "none") +
  scale_fill_manual(values = c("black", "red", "green", "blue", "cyan"))
```

```{r}
library(gridExtra)
grid.arrange(black_plot, white_plot, ncol = 2)
```


